cardinalR: Generating interesting high-dimensional data structures

### 2025-04-20 paper-cardinalR.Rmd
Updated second paragraph of intro

The R ecosystem offers a variety of packages for generating synthetic data, each with its strengths and focus. For instance, **`geozoo` (Schloerke, [Year of last commit on GitHub])** provides a rich collection of geometric objects, facilitating the creation and analysis of specific shapes, primarily in lower-dimensional spaces. **`snedata` (Melville, [Year of last commit on GitHub])** offers tools for generating simplified datasets tailored for evaluating dimensionality reduction techniques like t-SNE, often focusing on revealing low-dimensional embeddings of complex structures. Packages like **`mlbench` (Leisch & Dimitriadou, 2010)** provide a collection of well-known benchmark datasets, often with established classification or regression challenges. In the realm of single-cell omics, **`splatter` (Zappia et al., 2017)** excels at simulating complex biological data, capturing nuances like batch effects and differential expression.

>>>>

There are numerous packages available in R for generating synthetic data, each designed with unique characteristics and focus areas. For example, `geozoo` (@barret2016) offers a large collection of geometric objects, allowing users to create and analyze specific shapes, primarily in lower-dimensional spaces. Another useful package is `snedata` (@james2025), which provides tools for generating simplified datasets useful for evaluating dimensionality reduction techniques like tSNE, often focusing on understanding and evaluating low-dimensional embeddings of complex data structures. Additionally, `mlbench` (@friedrich2024) includes a collection of well-known benchmark datasets commonly associated with established classification or regression challenges. In the field of single-cell omics, `splatter` (@luke2017) is particularly simulate complex biological data, effectively capturing nuances such as batch effects and differential expression.
